---
title: "EDMS 646: Homework 3"
author: "Minoo Ahmadi"
date: "March 9, 2017"
output: pdf_document
header-includes: \usepackage{float}
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
```

```{r include=FALSE}
getwd()
setwd("C:/Users/ahmadi/Documents/Courses/EDMS646/EDMS646_Spring_2017/Hwk3")
hsb1.data <- read.table ("HSB1.csv", header = TRUE, na.strings="-99", sep = ",")
summary(hsb1.data)
head(hsb1.data)
tail(hsb1.data)
```



\begin{center}
\Large \textbf{PART 1: Multiple Regression - Initial model: Use data set HSB1}
\end{center}

1. 
```{r include=FALSE}
hsb1.lm <- lm(science~locus + concept + mot, data=hsb1.data)
summary(hsb1.lm)
names(hsb1.lm)
```
\[\hat{Y} = 50.3514 + 4.7188(\text{locus}) + 0.2549(\text{concept}) + 1.7113(\text{mot})\]

Controlling for self-concept and motivation, one unit increase in locus of control will increase the science score by 4.7188 units.
Controlling for locus of control and motivation, one unit increase in self-concept will increase the science score by 0.2549 units.
Controlling for locus of control and self-concept, one unit increase in motivation will increase the science score by 1.7113 units.

2.
H_0: $\beta_{locus} = \beta_{concept} = \beta_{mot} = 0$
H_1: $\beta_j \neq 0$



3.


4. 


\begin{center}
\Large \textbf{Part 2: Multiple Regression - Final model: Use data set HSB1}
\end{center}

1.


2. It looks evenly distributed to me and I don't seem to detect any outliers here.



\begin{center}
\Large \textbf{Part 3: ANOVA using Multiple Regression}
\end{center}

1.

2.

3.


