---
title: "EDMS 646: Homework 3"
author: "Minoo Ahmadi"
date: "March 9, 2017"
output: pdf_document
header-includes: \usepackage{float}
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
```

```{r include=FALSE}
getwd()
setwd("C:/Users/ahmadi/Documents/Courses/EDMS646/EDMS646_Spring_2017/Hwk2")
burt.data <- read.table ("Burt.csv", header = TRUE, sep = ",")
summary(burt.data)
head(burt.data)
tail(burt.data)
```



\begin{center}
\Large \textbf{PART 1: Multiple Regression - Initial model: Use data set HSB1}
\end{center}

1. There's a relatively strong positive linear relationship between IQ score of identical twins raised in foster homes (FostIQ) and that of their siblings whom were raised in the biological parent's homes (QwnIQ).
```{r include=FALSE}

library(ggplot2)
dev.off()
```

```{r echo=TRUE, fig.width= 3, fig.height= 3}
ggplot(burt.data, aes(OwnIQ, FostIQ)) + geom_point(size = 2) + geom_smooth(method = "lm") +
   theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(),
        axis.ticks = element_line(),
        
     )

```


2.
```{r}
cor.test(burt.data$OwnIQ, burt.data$FostIQ)

```
correlation coefficient (r): 0.8767131
p < .001

3.
\begin{align*}
t &= r \sqrt{\frac{n-2}{1-r^2}}= 0.8767131 \sqrt{\frac{51}{0.2313741}}= 13.01623\\
\nu &= n-2 = 53-2 = 51\\
\rightarrow t(51) &= 13.01
\end{align*}

4. The critical t-value is $\pm 2.008$ and the observed t-value exceeds this value. Hence, r is statistically significant at the $\alpha = .05$.





\begin{center}
\Large \textbf{Part 2: Multiple Regression - Final model: Use data set HSB1}
\end{center}

1.
```{r include=FALSE}
reg.data <- read.table ("REG1.csv", header = TRUE, sep = ",")
head(reg.data)

```


```{r}
reg.lm<- lm (Y~X, data = reg.data)
summary(reg.lm)
```

2. It looks evenly distributed to me and I don't seem to detect any outliers here.

```{r include = FALSE}
require(MASS)
require(plotrix)

st.residuals <- stdres(reg.lm)
```

```{r fig.width= 3.5, fig.height= 3.5}
plot(st.residuals, ylim = c(-3, 3))
```


\begin{center}
\Large \textbf{Part 3: ANOVA using Multiple Regression}
\end{center}

1.

\begin{table}[H]
\centering
\begin{tabular}{c|c|c|c|c|c}
Source & SS & df & MS & F & sig. \\\hline
Regression & 300 & 4 & 75 & 59.25 & <.001 \\
Residuals & 500 & 5 & 1.26\\
Total & 800 & 9
\end{tabular}
\caption{\label{tab:widgets}The ANOVA table.}
\end{table}


2.

3.


