---
title: "EDMS 646: Homework 3"
author: "Minoo Ahmadi"
date: "March 9, 2017"
output: pdf_document
header-includes: \usepackage{float}
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
```

```{r include=FALSE}
getwd()
# setwd("C:/Users/ahmadi/Documents/Courses/EDMS646/EDMS646_Spring_2017/Hwk3")
hsb1.data <- read.table ("HSB1.csv", header = TRUE, na.strings="-99", sep = ",")
summary(hsb1.data)
head(hsb1.data)
tail(hsb1.data)
```

```{r}
install.packages("psych", dependencies=TRUE)
library(lattice)

```

\begin{center}
\Large \textbf{PART 1: Multiple Regression - Initial model: Use data set HSB1}
\end{center}

1. 
```{r include=FALSE}

hsb1.lm <- lm(science~locus + concept + mot, data=hsb1.data)
summary(hsb1.lm)
names(hsb1.lm)
```
\[\hat{Y} = 50.3514 + 4.7188(\text{locus}) + 0.2549(\text{concept}) + 1.7113(\text{mot})\]

Controlling for self-concept and motivation, one unit increase in locus of control will increase the science score by 4.7188 units.
Controlling for locus of control and motivation, one unit increase in self-concept will increase the science score by 0.2549 units.
Controlling for locus of control and self-concept, one unit increase in motivation will increase the science score by 1.7113 units.


Independence means that, the Y values are statistically independentof one another.
Linearity means that, the mean value Y isproportional to the independent variable (X). Hence, a straight line function.
Normally Distributed means that, for a fixed value of X,Yhas a normal distribution.
Homoscedasticity or Homogeneity means that,the varianceofYis the same for any X (constant variance). 

```{r echo=TRUE, fig.width= 3, fig.height= 3}
ggplot(burt.data, aes(OwnIQ, FostIQ)) + geom_point(size = 2) + geom_smooth(method = "lm") +
   theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(),
        axis.ticks = element_line(),
        
     )

```


2.
```{r}
cor.test(burt.data$OwnIQ, burt.data$FostIQ)

```
correlation coefficient (r): 0.8767131
p < .001

3.
\begin{align*}
t &= r \sqrt{\frac{n-2}{1-r^2}}= 0.8767131 \sqrt{\frac{51}{0.2313741}}= 13.01623\\
\nu &= n-2 = 53-2 = 51\\
\rightarrow t(51) &= 13.01
\end{align*}

4. The critical t-value is $\pm 2.008$ and the observed t-value exceeds this value. Hence, r is statistically significant at the $\alpha = .05$.





\begin{center}
\Large \textbf{Part 2: Multiple Regression - Final model: Use data set HSB1}
\end{center}

1.
```{r include=FALSE}
reg.data <- read.table ("REG1.csv", header = TRUE, sep = ",")
head(reg.data)

```


```{r}
reg.lm<- lm (Y~X, data = reg.data)
summary(reg.lm)
```

2. It looks evenly distributed to me and I don't seem to detect any outliers here.

```{r include = FALSE}
require(MASS)
require(plotrix)

st.residuals <- stdres(reg.lm)
```

```{r fig.width= 3.5, fig.height= 3.5}
plot(st.residuals, ylim = c(-3, 3))
```


\begin{center}
\Large \textbf{Part 3: ANOVA using Multiple Regression}
\end{center}

1.

\begin{table}[H]
\centering
\begin{tabular}{c|c|c|c|c|c}
Source & SS & df & MS & F & sig. \\\hline
Regression & 300 & 4 & 75 & 59.25 & <.001 \\
Residuals & 500 & 5 & 1.26\\
Total & 800 & 9
\end{tabular}
\caption{\label{tab:widgets}The ANOVA table.}
\end{table}


2.

3.


