---
title: "EDMS 646: Homework 2"
author: "Minoo Ahmadi"
date: "February 23, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require("knitr")
opts_knit$set(root.dir = "~/gits/project/")
```

```{r}
getwd()
burt.data <- read.table ("Burt.csv", header = TRUE, sep = ",")
summary(burt.data)
head(burt.data)

```



\begin{center}
\Large \textbf{Part 1: Correlation}
\end{center}

\begin{enumerate}

\item 

\item 

\item 

\item How to write Mathematics

\LaTeX{} is great at typesetting mathematics. Let $X_1, X_2, \ldots, X_n$ be a sequence of independent and identically distributed random variables with $\text{E}[X_i] = \mu$ and $\text{Var}[X_i] = \sigma^2 < \infty$, and let
\[S_n = \frac{X_1 + X_2 + \cdots + X_n}{n}
      = \frac{1}{n}\sum_{i}^{n} X_i\]
denote their mean. Then as $n$ approaches infinity, the random variables $\sqrt{n}(S_n - \mu)$ converge in distribution to a normal $\mathcal{N}(0, \sigma^2)$.


\item

\end{enumerate}

\begin{center}
\Large \textbf{Part 2: Simple Regression}
\end{center}

\begin{enumerate}
\item
\item
\item blah blah
\begin{enumerate}
\item Like this,
\item and like this.
\item
\end{enumerate}
\item
\item
\item mhmb
\begin{enumerate}
\item 
\item
\item 
\item
\item 
\item
\item 
\item
\item
\end{enumerate}
\item geez
\begin{enumerate}
\item 
\item
\item
\item
\end{enumerate} 
\end{enumerate}

\begin{center}
\Large \textbf{Part 3: ANOVA Table}
\end{center}

\begin{enumerate}
\item 
\end{enumerate}



